% ============================================================
%  DP-LightGCN: Differentially Private LightGCN with 
%  Layer-Wise Adaptive Budget Allocation
% ============================================================
\documentclass[journal]{IEEEtran}
\usepackage{amsmath,amssymb,amsfonts}
\usepackage{booktabs}
\usepackage{graphicx}
\usepackage{hyperref}
\usepackage{cite}

\begin{document}

\title{DP-LightGCN: Differentially Private LightGCN with 
Layer-Wise Adaptive Budget Allocation for 
Privacy-Preserving Recommendation}

\author{Zhou Zhou{1}$}
\maketitle

\begin{abstract}
Graph Neural Networks (GNNs), particularly LightGCN, have achieved 
remarkable success in collaborative filtering-based recommendation 
systems. However, user-item interaction data raises significant 
privacy concerns. Existing differentially private recommendation 
methods primarily focus on matrix factorization and employ simplistic 
noise injection strategies. We propose DP-LightGCN, integrating 
embedding-level Gaussian perturbation with a novel layer-wise adaptive 
budget allocation mechanism. Our method dynamically allocates privacy 
budgets based on layer-wise embedding norms, providing better 
protection where needed. We evaluate on Gowalla and Amazon-Book, 
demonstrating that DP-LightGCN outperforms uniform and decreasing 
strategies by up to 56\% in Recall@20 under strong privacy 
($\varepsilon=5$).
\end{abstract}

\section{Introduction}
Recommendation systems are integral to modern platforms. 
Graph Neural Networks (GNNs), especially LightGCN~\cite{he2020lightgcn}, 
have achieved state-of-the-art performance. However, these models 
are vulnerable to privacy attacks.

Differential Privacy (DP)~\cite{dwork2006calibrating} is the 
standard for privacy protection. We propose DP-LightGCN with:
(1) layer-wise adaptive budget allocation,
(2) adaptive clipping, and
(3) embedding-level DP perturbation.

\section{Related Work}
McSherry and Mironov~\cite{mcsherry2009differentially} first 
introduced DP to recommendation. Friedman et al.~\cite{friedman2016differential} 
extended to matrix factorization. Jia et al.~\cite{jia2020towards} 
proposed DP-GCN. Li et al.~\cite{li2024p3gnn} proposed P3GNN.

\section{Preliminaries}
LightGCN~\cite{he2020lightgcn} propagates: $\mathbf{E}^{(k+1)} = \mathbf{A} \mathbf{E}^{(k)}$.

($\varepsilon,\delta$)-DP: $\Pr[\mathcal{M}(\mathcal{D}) \in \mathcal{S}] \leq e^{\varepsilon} \Pr[\mathcal{M}(\mathcal{D}') \in \mathcal{S}] + \delta$.

\section{Proposed Method}
Adaptive clipping:  = \max(2 \cdot \frac{1}{n} \sum \|\mathbf{e}_i^{(k)}\|_2, 0.1)$.

Noise scale: $\sigma_k = C_k \cdot \sqrt{2 \log(1.25/\delta)} / \varepsilon_k$.

Budget strategies: Uniform: $\varepsilon_k = \varepsilon/K$, 
Decreasing: $\varepsilon_k = \varepsilon \cdot (1/k)/\sum 1/j$, 
Adaptive: $\varepsilon_k = \varepsilon \cdot \|\mathbf{E}^{(k)}\|_F / \sum \|\mathbf{E}^{(j)}\|_F$.

\section{Experiments}
\begin{table}[h]
\centering
\caption{Results on Gowalla.}
\begin{tabular}{lcc}
\hline
Method & R@20 & N@20 \\
\hline
LightGCN (No DP) & 0.1064 & 0.0914 \\
DP-Uniform ($\varepsilon$=5) & 0.0288 & 0.0258 \\
DP-Decreasing ($\varepsilon$=5) & 0.0361 & 0.0329 \\
\textbf{DP-Adaptive ($\varepsilon$=5)} & \textbf{0.0448} & \textbf{0.0407} \\
\hline
\end{tabular}
\end{table}
Adaptive outperforms uniform by 56\% at $\varepsilon=5$.

\section{Conclusion}
We proposed DP-LightGCN with adaptive budget allocation.

\begin{thebibliography}{99}
\bibitem{he2020lightgcn} X. He et al., `LightGCN,'' SIGIR 2020.
\bibitem{dwork2006calibrating} C. Dwork et al., TCC 2006.
\bibitem{mcsherry2009differentially} F. McSherry et al., KDD 2009.
\bibitem{friedman2016differential} A. Friedman et al., JPC 2016.
\bibitem{jia2020towards} J. Jia et al., NeurIPS 2020.
\bibitem{li2024p3gnn} J. Li et al., KBS 2024.
\end{thebibliography}
\end{document}
